---
title: "CLEAR Trial Data Cleaning - DAILY"
author: "Jordan Barone"
date: "8/23/2021"
output: html_document
---

#markdown setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

#script setup
```{r}

#only run this line if you want to clear your workspace befor starting
rm(list=ls())

#load libraries
library(ggplot2) #plot
library(lubridate) #clean dates
library(knitr) #make easy tables
library(janitor) #some useful cleaning functions
library(stringr) #clean by matching strings
library(haven) #preferred import pacakge for SPSS files
library(tidyverse) #syntax
library(skimr) #some useful summary functions
library(visdat) #visualize data - great for missingness
library(assertive) #use for functions like "assert_all_are_in_past" 


#read in daily dataset
#recommended: download from qualtrics as an SPSS file, this allows us to preserve label options and values with nice formatting

#change the name of the object (left of "<-") and change your path to whatever dataset you're working with (right of "<-")
Clear2DailyData <- read_spss("~/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CLEAR2/02_datasets/CLEAR2_DAILY/01_raw_data/2022-07-28 clear2daily raw.sav")


```

#date prep and cleaning
-glimpse %>% summary: ensure that the number of total rows and columns seems good

-skim function: check that id and daterated have NO missing values
  - note on skim: it can't read labelled doubles (aka the type of variable SPSS contains, with labels embedded), so it will think everything is a character - not a problem
```{r}

## get overview of data ## 
glimpse(Clear2DailyData) %>% summarize() #get a row value, column value, variable type, and see the first few values of each variable
skim(Clear2DailyData) %>% View() #i like View when the df will be bigger than my console screen; this is a good function to check basic missingness, mins and maxs, etc

#check ID
str(Clear2DailyData$id) #see variables and labeled options
Clear2DailyData$id <- as.numeric(Clear2DailyData$id) #format as numeric

#check date
str(Clear2DailyData$daterated)
#format all CLEAR2 dates as R-readable dates (lubridate package)
Clear2DailyData$daterated <- mdy(Clear2DailyData$daterated)

#add weekend variable to clear2
# step 1: add dayofweek to clear2
## note that R lubridate defaults to 1=Monday, we want 1=Sunday to match CLEAR-1 SAS convention
## THIS CAN VARY based on how you want to set your week day start - just be sure you know
Clear2DailyData$dayofweek <- wday(Clear2DailyData$daterated, week_start = 7) 

# step 2: days 6/7/1 are friday/saturday/sunday
Clear2DailyData <- Clear2DailyData %>% 
  mutate(frisatsun = ifelse(dayofweek >=6 | dayofweek==1, 1, 0))
```

# check each dataset (how many rows? vars? participants?)
```{r interim data check}

##CLEAR2:
#number of total surveys
dim(Clear2DailyData)[1]
#number of variables
dim(Clear2DailyData)[2] 
#number of participants
length(unique(Clear2DailyData$id))
```

# Rvisualize data
```{r}

#visualize the variables types in your dataset
vis_dat(Clear2DailyData, warn_large_data = FALSE)

#visualize where you have missing data
vis_miss(Clear2DailyData, warn_large_data = FALSE)
```

#Example: clean and confirm ID
*the type of errors you may see that MUST be fixed:*
  # ",259" should be "259"
  # 2520 should be 252
  # 299 is not a participant
  # NA is not an ID, and you need to find out if this was a blank row, or if the data belongs to someone
```{r}

#are all the ID numbers correct? see notes in text above
Clear2DailyData %>% group_by(id) %>% summarize()

#is there a participant you KNOW you need to remove?
Clear2DailyData <- Clear2DailyData %>%
  filter(id !=228)

```

#row and column cleaning
``` {r}

#remove any rows or columns that are all NA
Clear2DailyData <- Clear2DailyData %>% remove_empty()

#how many rows now?
dim(Clear2DailyData) [1] 

#how many columns now?
dim(Clear2DailyData) [2]

#what are the column names?
colnames(Clear2DailyData)

#remove columns from qualtrics metadata
Clear2DailyData <- Clear2DailyData %>% select(-c(Status,
                                              Duration__in_seconds_,
                                              ResponseId,
                                              UserLanguage,
                                              DistributionChannel,
                                              IDN))

```


### export file [always make saving its own chunk so you remember to rename!]
```{r}
#CHANGE FILE NAME BELOW, then save!
write_sav(Clear2DailyData, "../03_cleaned_data/2022-07-28_clear2dailysurvey.sav")

##MAKE SURE YOU MOVE THE FILE INTO THE "CLEANED" FOLDER IF IT IS READY!

```