---
title: "model_which_ids_are_being_used_in_my_model?"
output: html_notebook
---

```{r libraries}
knitr::opts_chunk$set(echo = TRUE)
library(haven)
library(dplyr)
library(ggplot2)
library(kableExtra)
library(lubridate)
library(readr)
library(tidyverse)
library(psych)
library(lme4)
library(nlme)
library(gridExtra)
library(lmerTest) #using lmerTest to get p-values
library(dfoptim)
library(optimx)
library(interactions)
library(latex2exp)
library(emmeans)
library(plotly)
knitr::opts_knit$set(root.dir = "C:/Users/anish/Documents/Clear Lab Repositories/code_templates_AN") #change this to where your local repo's are stored 
```

### Why are the number of id's in my model different from lme4 or lmertest is using?

This is code to show you how to extract the id groups in your model and compare them to the id's in your analytic dataset. You can then investigate the id's that are not being used in your model and determine why they are not being used in the model (e.g. dropped out or withdrawn right after enrollment, did not have enough cycledays, ghosted, dropped out before they menstruated, etc.). You can also use this code to ensure all your demographic tables and descriptive analyses are done on the list of id's that are being utilized by your analytic models. 

```{r}
# Load in data that is being inputted into your model
dat <- read_sav( "C:/Users/anish/Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CLEARCOMBINED_projects/CLEAR23_STRAIN/02_code_dataedits_output/STRAIN23_merged_data_prepped.sav")

#run your model
model1_NA <- lmerTest::lmer(neg_affect ~ pretime_scale + posttime_scale + StressCT_z + age_z + SSRI + StressCT_z:pretime_scale + StressCT_z:posttime_scale + SSRI:pretime_scale + SSRI:posttime_scale + (1 + pretime_scale|id), data = dat, REML = TRUE)
summary(model1_NA) #converges, no issues 
```

```



```{r}
# Find the unique IDs in the model data
 
 model1_NA_id <- unique(model1_NA@flist[[1]])
 model1_NA_id

#  #201  202  203  204  205  206  208  209  210  211  212  213  214  215  216  217  220  221  222  223  224 
#  225  226  227  229  231  234  235  236  237  239  240  241  242  243  244  247  249  251  252  254  255 
# 256  257  258  259  260  261  3005 3011 3012 3016 3017 3019 3021 3023 3024 3025 3026 3028 3029 3030 3031
#   3032 3041 3043 3044 3045 3046 3047 3048 3049 3050 3051 3052 3055 3056 3057 3058 3059 3060 3062 3064 3066
#  3067 3069 3070 3072 3073 3074 3077 3078 3079 3080 3081 3082 3083 3084 3085 3088 3089 3090 3091 3092 3093
#  3094 3095 3096 3097 3098 3099 3100 3101 3102 3104 3105 3106 3107 3108 3109 3110 3111 3114 3117
 
 model1_NA_miss <- setdiff(dat$id, model1_NA_id)
#218 3002 3004 3006 3008 3015 3054 3061 3068 3103 3115 3119 3120
 
#For every id in model1_NA_miss, how many observations of non-NA neg_affect exist in the dataframe dat
# Create a subset of dat with non-missing and non-empty neg_affect values
dat_modNA1 <- dat %>% 
  filter(!is.na(neg_affect) & neg_affect != "")

# Create a data frame to store the results
result_df <- data.frame(id = numeric(),
                        num_obs = numeric())

# Loop over IDs in model1_NA_miss and count the number of non-missing numerical observations
for (id in model1_NA_miss) {
  num_obs <- dat_modNA1 %>% 
    filter(id == !!id) %>% 
    pull(neg_affect) %>% 
    as.numeric() %>% 
    sum(na.rm = TRUE)
  
  result_df <- bind_rows(result_df, data.frame(id = id, num_obs = num_obs))
}

# Print the results in a tibble
result_tbl <- as_tibble(result_df)
print(result_tbl)


#function to investigate different id's that are not included in the model 
drsp_cols <- names(dat)[grep("^DRSP", names(dat))]


idInv <- function(numid){
  dat_id <- subset(dat, id == numid, select = c("id","daterated", "neg_affect", drsp_cols))
  names(dat_id) <- paste0("dat_", numid, "_", names(dat_id))
  return(dat_id)
}

dat_218 <- idInv(218)
dat_218 #only 5 cycle days, withdrew from study shortly after enrolling
dat_3054 <- idInv(3054) #only 4 cycle days, withdrew from study shortly after enrolling
dat_3061 <- idInv(3061) #dropped out shortly after enrolling
dat_3068 <- idInv(3068) #92 rows of missing data for DRSP items- look into 
dat_3103 <- idInv(3103) #discontinued 
dat_3115 <- idInv(3115) #23 rows, all missing data- look into
dat_3119 <- idInv(3119) #10 rows all missing data 
dat_3120 <- idInv(3120) #8 rows, all missing data, recently enrolld 
```

