---
title: "plot_implied_val_interaction_plots_splines"
author: "Anisha Nagpal"
date: "`r Sys.Date()`"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(haven) #importing .sav files into R
library(ggplot2) #creating plot labels
library(lme4) #MLM models 
library(emmeans) #lib for extracting predicted model vals for different implied interaction vals
library(latex2exp) #lib for using laTex in ggplot labels 

#set your working directory (where your Github files are saved locally on your computer)
knitr::opts_knit$set(root.dir = "C:/Users/anish/Documents/Clear Lab Repositories/code_templates_AN")
#Import your dataset that your model is based on
dat <- read_sav( "C:/Users/anish/Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CLEARCOMBINED_projects/CLEAR23_STRAIN/02_code_dataedits_output/STRAIN23_merged_data_prepped.sav")
```


### Write your model using Lme4 

```{r}
model1a <- lme4::lmer(neg_affect ~ pretime_scale + posttime_scale + StressCT + age_z + SSRI + StressCT:pretime_scale + StressCT:posttime_scale + SSRI:pretime_scale + SSRI:posttime_scale + (1 + pretime_scale |id), data = dat, REML = TRUE)
summary(model1a)
```

### Specifying Model Estimates for pretime spline

- Here we extract the mean predicted values of your model for the pretime spline 

```{r}
d1NApre <- as.data.frame(emmip(model1a, StressCT ~ pretime_scale, at = list(pretime_scale = seq(-1, 0, by = (1/13)), StressCT = c(4, 44, 73) ), CIs =TRUE, plotit = FALSE)) 
#at = list(timespline = c([values of time spline]), moderator =c([values of continous mod that you want to plot]))
##these value of time spline are -13 to +13, with menses centered at 0, and then scaled, so divided by 13. Therefore the pretime spline ranges from -1 to 0
#CIs = confidence intervals 
d1NApre$time <- "pretime" #specify that these model estimates are for pretime spline, by adding a string "pretime"
d1NApre$posttime_scale = 0 #posttime vals should be 0 when plotting the pretime spline 

```

### Specifying Model Estimates for posttime spline

- Here we extract the mean predicted values of your model for the posttime spline 

```{r}
d1NApost <- as.data.frame(emmip(model1a, StressCT ~ posttime_scale, at = list(posttime_scale = seq(0,1, by = (1/13)), StressCT = c(4, 44, 73) ), CIs =TRUE, plotit = FALSE)) 
#at = list(timespline = c([values of time spline]), moderator =c([values of continous mod that you want to plot]))
##these value of time spline are -13 to +13, with menses centered at 0, and then scaled, so divided by 13. Therefore the posttime spline ranges from 0 to 1
#CIs = confidence intervals 
d1NApost$time <- "posttime"
d1NApost$pretime_scale = 0 #pretime vals should be 0 when plotting the posttime spline 
```

### Combine Pretime and Postime estimates into 1 dataframe 

```{r}
d1NA <- rbind(d1NApre, d1NApost)
```

### InterPlot (plotting Function I wrote to follow lab conventions for counting cycle days) 

```{r}
interPlot <- function(d, mod, ylab, legend){
  ggplot(d, aes(x = xvar, y = yvar, group = factor(mod), color = factor(mod))) + 
    geom_line() +
    geom_point() + 
    xlab("Cycle Day") +
    ylab(ylab) + 
    labs(color = legend)+
    scale_x_continuous(limits = c(-1,1), breaks = round(seq(-1 ,1, by = (1/13)), 2), labels=c("-13", "-12", "-11", "-10", "-9", "-8", "-7", "-6", "-5", "-4", "-3", "-2", "-1", "1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12", "13", "14"))+
    geom_vline(xintercept = 0, color = "red") +
    geom_errorbar(aes(ymin = LCL, ymax = UCL), alpha = 0.2) +
    theme_bw()
}

```

### Plot!

```{r fig.height = 4, fig.width = 10}
interPlot(d1NA, d1NA$StressCT, "Negative Affect", "Number of Lifetime Stressors")
#parameters: combined dataframe of pretime and posttime estimates, moderator from dataframe of pretime and posttime estimates, xlab name, legend title 
```









