---
title: "Raincloud Plot Template Code"
author: "Jordan Barone"
date: '2022-06-24'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(ggdist)
library(ggforce)
```

Code based on: https://www.cedricscherer.com/2021/06/06/visualizing-distributions-with-raincloud-plots-and-how-to-create-them-with-ggplot2/

#Pull out your random effects by ID
-for my models, I ran pairwise contrasts between cycle phases and rotated the reference group. I have 3 versions of each model to get all pairwise combinations of periovulatory-perimenstrual-midfollicular-midluteal.
-model naming convention is modelX.Y.Z where
--> *x=outcome*
--> *y=model variation* (i.e., main effects only, interaction, with/without random effects - I tried several versions to get best convergence)
--> *z=reference group*
-so, to get all random effects for all pairwise comparisons of my main effects model with drinking as an outcome, I have to pull from 3 models with the same specifications, but rotating reference groups
``` {r}
#save out random effects and ID from your model
RRdata_randomeffects1.1.ov <- as.data.frame(ranef(RR_model1.1.ov)[["id"]])
#when the model is saved as a data frame, the IDs are assigned as the rownames - make ID a column
RRdata_randomeffects1.1.ov <- cbind(id = rownames(RRdata_randomeffects1.1.ov), 
                                    RRdata_randomeffects1.1.ov)
#since you just removed the incorrect rownames, make the new rownames just a regular sequence
rownames(RRdata_randomeffects1.1.ov) <- 1:nrow(RRdata_randomeffects1.1.ov)

#repeat above steps with reference group #2
RRdata_randomeffects1.1.mens <- as.data.frame(ranef(RR_model1.1.mens)[["id"]])
RRdata_randomeffects1.1.mens <- cbind(id = rownames(RRdata_randomeffects1.1.mens), 
                                      RRdata_randomeffects1.1.mens)
rownames(RRdata_randomeffects1.1.mens) <- 1:nrow(RRdata_randomeffects1.1.mens)
#repeat above steps with reference group #3
RRdata_randomeffects1.1.lut <- as.data.frame(ranef(RR_model1.1.lut)[["id"]])
RRdata_randomeffects1.1.lut <- cbind(id = rownames(RRdata_randomeffects1.1.lut), 
                                      RRdata_randomeffects1.1.lut)
rownames(RRdata_randomeffects1.1.lut) <- 1:nrow(RRdata_randomeffects1.1.lut)
```

#Step 2. Clean your random effects dataset
```{r}
#removing duplicates, renaming, adding fixed effects
RRdata_randomeffects1.1.ov <- RRdata_randomeffects1.1.ov %>%
  select(-c("(Intercept)")) %>% #remove the intercept column, I won't be using it
  #rename the predictor variables to denote what their reference group is
  rename(periov_v_midlut = LH_midluteal_count,
         periov_v_perimens = LH_perimenstrual_count,
         periov_v_midfol = LH_midfol_LH) %>%
  #MANUALLY add the fixed effect from the model to each random effect as a new variable
  mutate(periov_v_midfol_plusfixed = periov_v_midfol-0.09906,
         periov_v_perimens_plusfixed = periov_v_perimens-0.14326,
         periov_v_midlut_plusfixed = periov_v_midlut-0.23305)

#repeat above with reference group #2
#--> note here that the duplicate contrast we are removing is because already got "periov vs perimens" with periov as reference group, so we don't need "perimens vs periov"
RRdata_randomeffects1.1.mens <- RRdata_randomeffects1.1.mens %>% 
  select(-c(LH_periov_LH, "(Intercept)")) %>%
  rename(perimens_v_midlut = LH_midluteal_count,
         perimens_v_midfol = LH_midfol_LH) %>%
  mutate(perimens_v_midlut_plusfixed = perimens_v_midlut+0.37630,
         perimens_v_midfol_plusfixed = perimens_v_midfol+0.04419)

#repeat above with reference group #3
RRdata_randomeffects1.1.lut <- RRdata_randomeffects1.1.lut %>% 
  select(-c(LH_perimenstrual_count, LH_periov_LH, "(Intercept)")) %>%
  rename(midlut_v_midfol = LH_midfol_LH) %>%
  mutate(midlut_v_midfol_plusfixed = midlut_v_midfol-0.33211)

#merge to one random effect dataset
#(I just learned that the function "reduce" can merge multiple data frames, but this code goes two at a time)
RE1.1.ov.mens <- full_join(RRdata_randomeffects1.1.ov, RRdata_randomeffects1.1.mens, by="id")
RE1.1.ov.mens.lut <- full_join(RE1.1.ov.mens, RRdata_randomeffects1.1.lut, by="id")
```

#Step 3. Plot code
```{r}

RRPLOT3_randomeffects_model1.1 <- RE1.1.ov.mens.lut %>%
  #pivoting is necessary to get ALL random effects as one variable, rather than 6 columns
  pivot_longer(
    cols = ends_with("plusfixed"),
    names_to = "pairwise",
    values_to = "randomeffect") %>%
  #aesthetics will be your new pivoted columns
  ggplot(aes(x=pairwise, y=randomeffect, color=pairwise))+
  #boxplot!
  geom_boxplot(fill=NA,
                width = 0.2,
                outlier.color = NA)+
  #half histogram1
  ggdist::stat_halfeye(aes(fill=pairwise), 
                       adjust = 0.5, width=0.6, justification=-0.3, .width = 0, point_colour=NA)+
  #scatter!
  geom_point(size=1,alpha=0.2, position = position_jitter(seed=1, width = 0.1))+
  #everything from here down is just formatting and labeling
  scale_color_discrete(guide="none")+
  scale_fill_discrete(name="Cycle Phase Contrast",
                      labels=c("MIDLUT vs MIDFOL",
                               "PERIMENS vs MIDFOL",
                               "PERIMENS vs MIDLUT",
                               "PERIOV vs MIDFOL",
                               "PERIOV vs MIDLUT",
                               "PERIOV vs PERIMENS"))+
  coord_cartesian(xlim=c(1.2, NA), clip = "off")+
  theme_bw()+
  theme(axis.text.x = element_blank(),
        axis.title.x = element_blank())+
  ylab("Random Slope")+
  ggtitle("Random Effects",
          subtitle = "individual differences of cycle phase on drinking probability")+
  labs(fill="Cycle Phase Contrast")+
  geom_hline(yintercept = 0, linetype="dashed", color="gray")+
  geom_text(aes(x=1, y=-2.6), check_overlap = TRUE, fontface="italic", size=3,color="black",
            label="  (-)favors MIDLUT")+
  geom_text(aes(x=1, y=1.5), check_overlap = TRUE, fontface="italic", size=3,color="black",
            label="    (+)favors MIDFOL")+
  geom_text(aes(x=2, y=-1), check_overlap = TRUE, fontface="italic", size=3,color="black",
            label="(-)favors PERIMENS")+
  geom_text(aes(x=2, y=1), check_overlap = TRUE, fontface="italic", size=3,color="black",
            label="(+)favors MIDFOL")+
  geom_text(aes(x=3, y=-1.25), check_overlap = TRUE, fontface="italic", size=3,color="black",
            label="(-)favors PERIMENS")+
  geom_text(aes(x=3, y=2), check_overlap = TRUE, fontface="italic", size=3,color="black",
            label="(+)favors MIDLUT")+
  geom_text(aes(x=4, y=-0.75), check_overlap = TRUE, fontface="italic", size=3,color="black",
            label="(-)favors PERIOV")+
  geom_text(aes(x=4, y=0.5), check_overlap = TRUE, fontface="italic", size=3,color="black",
            label="(+)favors MIDFOL")+
  geom_text(aes(x=5, y=-2.25), check_overlap = TRUE, fontface="italic", size=3,color="black",
            label="(-)favors PERIOV")+
  geom_text(aes(x=5, y=2), check_overlap = TRUE, fontface="italic", size=3,color="black",
            label="(+)favors MIDLUT")+
  geom_text(aes(x=6, y=-1), check_overlap = TRUE, fontface="italic", size=3,color="black",
            label="(-)favors PERIOV")+
  geom_text(aes(x=6, y=0.75), check_overlap = TRUE, fontface="italic", size=3,color="black",
            label="(+)favors PERIMENS")
RRPLOT3_randomeffects_model1.1 %>% ggsave(filename = "2022-06-19 RR ranef_dd_pairwise.png", 
                                                width = 9, height = 3,
                                                path = "../04_writing/2022-04-26 ACER R&R/",
                                                units = c("in"),
                                                device = "png")
```