---
title: "person means devs and rolling avgs"
author: "Jordan Barone"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Import Data

Use this first code chunk below to import your libraries, set your working directory (file path where you would like to save your output), import your data, and subset your data. The libraries included below are some commonly useful libraries. Using hashtags to add headers to your RMarkdown (outside of code chunks) is helpful for people using your file to navigate the file using the outline bar in Rstudio. Also you can label all of your code chunks to help with file navigation. Here is a cheatsheet for RMarkdown: https://www.rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf

```{r libraries and data import}
library(haven) #for reading files, especially useful for importing SPSS files 
library(ggplot2) #for plotting
library(tidyverse) #for data management and manipulation


#setwd("")

#read_csv() #using haven, you can also use read.csv() in base R

#make sure using csv with correct sub-sample 


```

## FUNCTIONS TO CALCULATE PERSON CENTERED MEANS AND DAILY DEVIATIONS
create.person.mean is a function that takes your dataset (df), your variable (var), and any grouping variable (will usually be *id*) to create group-based means.
- new variable gets added to the dataset with the suffix ".m" after the original variable name

create.deviation is a function that your dataset (df), your variable (var), and the person-mean you already made (var.m) to create a daily deviation for that variable on that day. 
- new variable gets added to the dataset with the suffix ".d" after the original variable name
```{r}
#function to create person means
create.person.mean <- function(df, var, ...) {
  df %>%
    group_by(...) %>%
    mutate("{{var}}.m" := mean({{var}}, na.rm=T))
}
#function to create person deviations (note, must have person means already made)
create.deviation <- function(df, var, var.m) {
  df <- df %>%
    rowwise() %>%
    mutate("{{var}}.d" := {{var}} - {{var.m}})
}


```

### create outcome lists
- you can run the above functions on any single variable. However, if you have many variables to run the function on, it's best to make a list of outcomes, and then loop the function over all of them.
```{r}

#create list of outcomes
outcomes <- c("SI_Mean",
              "passiveSI",
              "activeSI",
              "DRSP1_depblue",
              "DRSP2_hopeless",
              "DRSP6_rejsens",
              "DRSP16_overwhelm",
              "DRSP4_anxious",
              "DRSP10_diffconc",
              "DRSP7_angirr",
              "DRSP11_lethtired",
              "DRSP12_13_appoverate", 
              "DRSP14_sleptmore",
              "DRSP15_troublesleep",
              "DRSP17_outofcontrol",
              "DRSP3_worthguilt",
              "DRSP9_lessintmot",
              "DRSP5_moodswings",
              "DRSP8_intconflict", 
              "DRSP20_headache") %>% noquote()

#make a list of deviation outcomes (this is useful if you are going to create rolling averages)
outcomes.d <- outcomes %>% paste0(".d") %>% noquote()
outcomes.d.roll <- outcomes.d %>% paste0(".roll") %>% noquote()

```


## execute functions
# to get a list of all person means: 
paste0(outcomelist,".m")
# to run create.deviation() one at a time.
c2d <- create.deviation(c2d, wishsleep, wishsleep.m)
# to run create.person.mean() one at a time.
 c2d <- create.person.mean(c2d, ASIQ4_thoughtwhen, id)
```{r}
#execute for loop: run create.person.mean() on everything in "outcomes"
for (i in outcomes) {
  PPc2d <- create.person.mean(PPc2d, !!sym({{i}}), id)
}
#execute for loop: run create.deviation() on everything in list
for (i in outcomes) {
  PPc2d <- create.deviation(PPc2d, !!sym({{i}}), !!sym(paste0({{i}}, ".m")))
}
```

#create rolling avgs

```{r}

create.rolling.deviation <- function(df, var.d) {
  df <- df %>%
    group_by(id) %>%
    mutate("{{var.d}}.roll" := rollapply({{var.d}}, 3, mean, align = "center", fill = NA)) %>%
    ungroup()
}
  

    
for (i in outcomes.d) {
  PPc2d <- create.rolling.deviation(PPc2d, !!sym({{i}}))
}

# next step here: work on fill arguments to change missingness to interpolate with avg

```
