---
title: "2022-07-12 Cycle Phase Coding for KY-ADHD"
author: "Jordan Barone"
date: '2022-07-12'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Step 1: packages, working directory, import data
```{r setup}

#libraries: not all of these will be used in this script, but this list is pretty exhaustive for data cleaning
library(haven) #for reading files
library(ggplot2) #for plotting
library(tidyverse) #for data management and manipulation
library(zoo) #for rolling averages
library(janitor) #nice cleaning functions
library(psych) #good describe/summary functions
library(visdat) #visualize data
library(skimr) #useful skimming functions for big datasets
library(lubridate) #for manipulating dates
library(readxl) #if you are importing an excel with multiple sheets

#set your working directory (this is where the code and any output will get saved)
setwd("~/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CLEAR2/02_datasets/CLEAR2_DAILY/02_data_prep_workspace/")

#import your file
#recommended for CLEAR Lab: use SPSS file format downloaded from qualtrics - this allows us to preserve label options and values with nice formatting
Clear2DailyData <- read_spss("~/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/CLEAR2/02_datasets/CLEAR2_DAILY/01_raw_data/2022-04-15 clear2daily raw.sav")

```


#Step 2: basic data review - make notes of what might need upstream or in-code cleaning
**skim function**: check that id and daterated have NO missing values
  - absolutely necessary to not have missing ID or DATES for the cycle day sequence code to be accurate!!!
  - note on skim: it can't read labelled doubles (aka the type of variable SPSS contains, with labels embedded), so it will think everything is a character - not a problem
```{r review your dataset}

## get overview of data ## 
glimpse(Clear2DailyData) %>% summarize() #get a row value, column value, variable type, and see the first few values of each variable
skim(Clear2DailyData) %>% View() #i like View when the df will be bigger than my console screen; this is a good function to check basic missingness, mins and maxs, etc

#visualize the types of variables included in your dataset, review any that seem miscoded
vis_dat(Clear2DailyData, warn_large_data = FALSE)
#visualize missing data, pay specific attention to any missing ID or date
vis_miss(Clear2DailyData, warn_large_data = FALSE)

#number of rows
dim(Clear2DailyData) [1]
#number of columns
dim(Clear2DailyData) [2]

#number of participants
length(unique(Clear2DailyData$id))


```


#Step 3: basic data cleaning
**date formatting**: the specific function you use will depend on how participants entered the date into the survey. This script uses mdy() to take data that was entered as mm/dd/yyyy and force it into a date-format. Will vary if data entry was different, look into "lubridate" package for more info.
```{r clean dataset}

#remove any entirely empty rows or columns
Clear2DailyData <- Clear2DailyData %>% remove_empty()

#check ID
str(Clear2DailyData$id) #see variables and labeled options
#make sure ID is formatted as a numeric variable
Clear2DailyData$id <- as.numeric(Clear2DailyData$id)
#make sure no "empty" rows got included - for our purposes, this would be any row that doesnt have an ID entered
Clear2DailyData <- Clear2DailyData %>%
  filter(!is.na(id))

#check date
str(Clear2DailyData$daterated)
#format all CLEAR2 dates as R-readable dates (mdy function lubridate package)
Clear2DailyData$daterated <- mdy(Clear2DailyData$daterated)

Clear2DailyData <- Clear2DailyData %>%
  filter(!is.na(id))

# filter out based on IDs (if using LH-based coding, you want to only include participants who have had a positive LH test, otherwise the code will be weird and give you a million zeroes)

Clear2DailyData <- Clear2DailyData %>%
  filter(id %in% LHids)

```


### step 4a: add cycleday, based on menses onset (menses onset = day 1)
```{r cycleday}

#if menstrualbleeding = 0.5 (spotting) or 1 (bleeding) and firstdayofperiod = 1, then cycleday = 1
Clear2DailyData$cycleday <- ifelse(Clear2DailyData$menstrualbleeding >= 0.5 
                                   & Clear2DailyData$firstdayofperiod == 1, 1, 0)
#convert all NAs in cycleday to 0
Clear2DailyData$cycleday[is.na(Clear2DailyData$cycleday)] <- 0

#Fill in missing days by ID 
Clear2DailyData <- Clear2DailyData[!is.na(Clear2DailyData$daterated), ]
Clear2DailyData <- Clear2DailyData %>% group_by(id) %>%
  complete(daterated = seq.Date(min(daterated), max(daterated), by = "day"))
Clear2DailyData <- cbind(Clear2DailyData, 
                         year = year(Clear2DailyData$daterated), 
                         month = month(Clear2DailyData$daterated), 
                         day = day(Clear2DailyData$daterated)) 
Clear2DailyData <- Clear2DailyData[with(Clear2DailyData,order(as.numeric(id), 
                                                              year, month, day, daterated), cycleday),]
Clear2DailyData$cycleday[is.na(Clear2DailyData$cycleday)] <- 0

Clear2DailyData <- Clear2DailyData[with(Clear2DailyData,order(id, 
                                                              year, 
                                                              month, 
                                                              day, 
                                                              daterated),cycleday),]

#make A the variable where menstrualbleeding starts AND period starts
Clear2DailyData$A <- Clear2DailyData$cycleday

cycleCount <- function(x) {
  #Get the index of 1
  inds <- which(x == 1)
  if(!length(inds)) return(0) 
  #Create a sequence with that index as 0
  num <- lapply(inds, function(i) {
    num <- seq_along(x) - i
    #Add 1 to values greater than equal to 0
    num[num >= 0] <- num[num >= 0] + 1
    num[num < -15 | num > 10] <- NA 
    num
  })
  #Select the first non-NA values from the sequence
  do.call(coalesce, num)
}

Clear2DailyData <- Clear2DailyData %>% group_by(id) %>% 
  mutate(cycleday = cycleCount(A))
Clear2DailyData <- Clear2DailyData %>% select(!A)

```

#review step 4a: 
```{r cycleday checking}

#how many daily ratings got assigned each cycle day?
Clear2DailyData %>% group_by(cycleday) %>% summarize(n=n())

#did any cycledays get assigned 0? this would only occur if the person never reported menses
#flag this ID for review and fix prior to adding phasing
Clear2DailyData %>% filter(cycleday==0) %>% pull(id)
```

#step 4b: make daycountLH, based on positive LH test (LH test day = 0)
```{r daycountLH}
#make A a new temp column where pt received pos ov test
Clear2DailyData$A <- Clear2DailyData$posLHtoday

#FUNCTION for calculating the sequence
LHCount <- function(x) {
  #Get the index of 1
  inds <- which(x == 1)
  if(!length(inds)) return(0) 
  #Create a sequence with that index as 0
  num <- lapply(inds, function(i) {
    num <- seq_along(x) - i
    num[num < -7 | num > 15] <- NA 
    num
  })
  #Select the first non-NA values from the sequence
  do.call(coalesce, num)
}

#run the LHCount function and save it as as a new column called daycountLH
Clear2DailyData<- Clear2DailyData %>% group_by(id) %>% 
  mutate(daycountLH = LHCount(A))
#remove the temp column A
Clear2DailyData <- Clear2DailyData %>% select(!A)
```

#review step 4b:
```{r daycountLH checking}

#how many daily ratings got assigned each cycle day?
Clear2DailyData %>% group_by(daycountLH) %>% summarize(n=n())

```

#step 5a: cycle phase based on menses count:
reminder: dummy code comes from Schmalenberger 2021 appendix
```{r cyclephase_count}

Clear2DailyData <- Clear2DailyData %>% 
  mutate(midluteal_count = ifelse(cycleday >= -9 & cycleday <= -5, 1, 0),
         perimenstrual_count = ifelse(cycleday >=-3 & cycleday <=2, 1, 0),
         midfol_count = ifelse(cycleday >=4 & cycleday <=7, 1, 0),
         periov_count = ifelse(cycleday >= -15 & cycleday <= -12, 1, 0))

#add a non-dummy coded variable with all of cycle phase info above, which will be useful for future visualizations and summaries
Clear2DailyData <- Clear2DailyData %>% mutate(cyclephase_count = case_when(periov_count==1 ~ 1,
                                                     midluteal_count==1 ~ 2,
                                                     perimenstrual_count==1 ~ 3,
                                                     midfol_count==1 ~ 4,
                                                     TRUE ~ as.numeric(NA)))

```

#step 5b: cycle phase based on LH count:
-note: the perimenstrual phase is still coded based on menses count variable, but for the purpose of keeping the dummy code labeling clear, we name it *perimenstrual_LH* to indicate that this variable gets included in the dummy code set of LH-based phasing
-again, all of this information is from Schmalenberger 2021 appendix
```{r cyclephase_LH}

Clear2DailyData <- Clear2DailyData %>% 
  mutate(midluteal_LH = ifelse(daycountLH >= 6 & daycountLH <= 10, 1, 0),
         perimenstrual_LH = ifelse(cycleday >=-3 & cycleday <=2, 1, 0), #note perimenstrual is based on cycleday
         midfol_LH = ifelse(daycountLH >=-7 & daycountLH <=-3, 1, 0),
         periov_LH = ifelse(daycountLH >= -2 & daycountLH <= 1, 1, 0),
         earlylut_LH = ifelse(daycountLH >= 2 & daycountLH <= 5, 1, 0))

#fill out the dummy code, so that if any phase is 1, all others should be 0 (not NA)
Clear2DailyData <- Clear2DailyData %>%
  rowwise() %>%
  mutate(sumdummy = sum(midfol_LH,
                        periov_LH,
                        earlylut_LH,
                        perimenstrual_LH,
                        midluteal_LH, na.rm = T)) %>% 
  #if all phases=0, make the midluteal_LH variable NA instead of 0,
  #if perimenstrual by menses count = 1, change midluteal_LH to 0 instead of NA to fill out structural set,
  #otherwise, keep as is
  mutate(midluteal_LH = case_when(sumdummy==0 ~ as.numeric(NA),
                                  sumdummy==1 & perimenstrual_LH==1 ~ 0,
                                  TRUE ~ midluteal_LH), 
         #same for midfol
         midfol_LH = case_when(sumdummy==0 ~ as.numeric(NA),
                               sumdummy==1 & perimenstrual_LH==1 ~ 0,
                               TRUE ~ midfol_LH), 
          #same for periov
         periov_LH = case_when(sumdummy==0 ~ as.numeric(NA),
                               sumdummy==1 & perimenstrual_LH==1 ~ 0,
                               TRUE ~ periov_LH),
         earlylut_LH = case_when(sumdummy==0 ~ as.numeric(NA),
                                  sumdummy==1 & perimenstrual_LH==1 ~ 0,
                                  TRUE ~ earlylut_LH), 
         #if any other phase is 1, fill out perimenstrual to be 0
         perimenstrual_LH = case_when(sumdummy==0 ~ as.numeric(NA), 
                                        sumdummy==1 & midfol_LH==1 ~ 0,
                                        sumdummy==1 & periov_LH==1 ~ 0,
                                      sumdummy==1 & earlylut_LH==1 ~ 0,
                                        sumdummy==1 & midluteal_LH==1 ~ 0,
                                        TRUE ~ perimenstrual_LH))

#add a non-dummy coded variable with all of cycle phase info above, which will be useful for future visualizations and summaries
Clear2DailyData <- Clear2DailyData %>% mutate(cyclephase_LH = case_when(periov_LH==1 ~ 1,
                                                                        earlylut_LH==1 ~ 2,
                                                     midluteal_LH==1 ~ 3,
                                                     perimenstrual_LH==1 ~ 4,
                                                     midfol_LH==1 ~ 5,
                                                     TRUE ~ as.numeric(NA)))

```

#step 6: review for any days that got double-counted
-when using LH-based phasing for all 4 phases, the most likely possibility for overlap is a day that got labeled midfollicular by LH-count and perimenstrual by menses count. *Perimenstrual_count trumps midfollicular by LH count* in this instance.
-any other overlaps within the dummy code are likely miscoded or an abnormal cycle.
```{r}

#check overlaps: goal is to only have 6 possible group combinations here (1 row per dummy coded phase, then a row where ALL are NA)
Clear2DailyData %>%
  group_by(periov_LH, earlylut_LH, midluteal_LH, perimenstrual_LH, midfol_LH) %>%
  summarize(n=n())
#make note of errors


#example code to review midfol_LH and perimenstrual_LH overlaps, if they exist
Clear2DailyData %>%
  filter(midfol_LH==1) %>%
  filter(perimenstrual_LH==1) %>%
  select(id, daterated, posLHtoday, daycountLH,
         menstrualbleeding,
         firstdayofperiod, cycleday,
         midfol_LH, periov_LH, earlylut_LH,
         midluteal_LH, perimenstrual_LH) %>%
  View()
#in CLEAR studies, we are looking to see if the day in question is at the early end of the follicular phase as estimated by counting the days prior to ovulation
#example: if the day in question is -7 by LH count and +2 by menses count, this was a short follicular phase, and is probably OK to be included, but needs to be recoded as "perimenstrual" (see below)


```

#step 7: edit days that got double-counted
```{r}

Clear2DailyData  <- Clear2DailyData  %>%
  #if perimenstrual by firstdayofperiod and midfol by pre-LH count are both 1, perimen wins, make midfol=0
  mutate(midfol_LH = case_when(perimenstrual_LH==1 ~ 0,
                                        TRUE ~ midfol_LH)) 

#review midlut and perimens overlaps - can override midluteal with perimens IF luteal phase is short but still seems normal (such as 9-10 days)
## make notes of any SUPER short luteal phases (like 4-5 days) that you might want to flag for anovulation
Clear2DailyData %>% filter(midluteal_LH==1 & perimenstrual_LH==1) %>% 
  select(id, daterated, daycountLH, cycleday, midluteal_LH, perimenstrual_LH)

#code for if you are ready to override midluteals with perimens
Clear2DailyData  <- Clear2DailyData  %>%
  #if perimenstrual by firstdayofperiod and midlut by count after LH test are both 1, perimen wins, make midlut=0
  mutate(midlut_LH = case_when(perimenstrual_LH==1 ~ 0,
                                        TRUE ~ midlut_LH)) 

## ANY OTHER COMBOS need hand review!!

#good practice to review again and ensure that dummy code is fixed
Clear2DailyData %>%
  group_by(periov_LH, earlylut_LH, midluteal_LH, perimenstrual_LH, midfol_LH) %>%
  summarize(n=n())

```

#optional step 8: save dataset 
```{r}

Clear2DailyData %>% write_sav("2022-07-12 DATASET with cycle coding.sav")
#or#
Clear2DailyData %>% write.csv("2022-07-12 DATASET with cycle coding.csv")

```



