---
title: "Untitled"
author: "Jordan Barone"
date: "5/12/2022"
output: html_document
---

*Rules before using this template*
- make sure all data has been cleaned, ESPECIALLY date cleaning and removing date duplicates
- make sure dummy rows have already been inserted on missing days
- cycleday based on menses count alone must be completed already
- phase-meds-tx tracking sheet must be cleaned and merged into your daily dataset

#override the participant report of positive ovulation from qualtrics with our manual, staff-entered ovulation from phase-meds-tx tracking
```{r}

#select the BEST lh variable for merging
c2daily <- c2daily %>% select(-posLHtoday) #removes the qualtrics version of the variable
c2daily <- c2daily %>% 
  rename(posLHtoday = posovtest) #renames the staff-entered variable to be "posLHtoday" bc that matches other codes (not a required step, but jb did this for merging with clear1, relevant for alcohol paper)
```

#some quick descriptives before diving in
```{r}

#participants who have at least 1 positive ov test
c2daily %>% group_by(id) %>% 
  filter(posLHtoday==1) %>% 
  summarize(n=n())

#make list of IDs with LH-confirmed ovulatory cycles
c2_LHids <- c2daily %>%
  filter(posLHtoday==1) %>% 
  group_by(id) %>%
  distinct(id) %>%
  select(id) %>%
  as.matrix() %>%
  c()

#can plug in any variable of interest here to find out how many participants in the LH-confirmed sample also met another criteria
#example: how many people in the LH-confirmed sample also had at least one instance of binge drinking?
c2daily %>% 
  filter(id %in% c2_LHids) %>%
  group_by(id, binge_today) %>%
  summarize(n=n()) %>%
  filter(binge_today==1)

#filter your group of just people with confirmed LH ovulation
c2daily_LHsample <- c2daily %>%
  filter(id %in% c2_LHids)

```

#create daycountLH: count from -7 to +10 surrounding ovulation as day 0
```{r}

#make A a new temp column where pt received pos ov test
c2daily_LHsample$A <- c2daily_LHsample$posLHtoday

#FUNCTION for calculating the sequence
LHCount <- function(x) {
  #Get the index of 1
  inds <- which(x == 1)
  if(!length(inds)) return(0) 
  #Create a sequence with that index as 0
  num <- lapply(inds, function(i) {
    num <- seq_along(x) - i
    num[num < -7 | num > 15] <- NA 
    num
  })
  #Select the first non-NA values from the sequence
  do.call(coalesce, num)
}

#run the LHCount function and save it as as a new column called daycountLH
c2daily_LHsample <- c2daily_LHsample %>% group_by(id) %>% 
  mutate(daycountLH = LHCount(A))
#remove the temp column A
c2daily_LHsample <- c2daily_LHsample %>% select(!A)

```


#create LH-based cycle phase, FOR BASELINE DAYS ONLY
***this process is different if you are including the experimental days***
The experimental condition starts at day +7. Using LH-surge methods, the midluteal phase would fall on days +6 to +10. Days +7 to +10 need to be excluded from any baseline-only papers (Barone Alcohol Paper, Ross SUI Mediators Paper). That gives us a max luteal phase of 1 day only - so we will use the backward count method for midluteal phase. cycleday (-15 to +10) and counting-based cycle phases (see HTSMC appendix) are already saved in the dataset that was imported.
*JBs naming conventions:* 
--> LHgroup_ prefix to indicate this counting system  (structural set of dummy codes, added LHgroup_ prefix per Katja paper) only applies to the sample with at least 1 confirmed ovulation
--> _LH/_count suffix to indicate if this phase is getting coded by its relation to LH-surge, or its relation to menses onset
*Rules:*
LHgroup_midfol_LH = day -7 to -3 before posLHtoday
LHgroup_periov_LH = day -2 to +1 surrounding posLHtoday
LHgroup_midluteal_count = day -9 to -5 before mensesonset
LHgroup_perimenstrual_count = day -3 to +2 surrounding menses onset

```{r}

#create midfol and periov phases via LH method, and copy midlut and perimens from counting methods
c2daily_LHsample <- c2daily_LHsample %>%
  group_by(id) %>%
  mutate(LHgroup_midfol_LH = ifelse(daycountLH>=-7 & daycountLH<=-3, 1, 0),
         LHgroup_periov_LH = ifelse(daycountLH>=-2 & daycountLH<=1, 1, 0),
         LHgroup_earlylut_LH = ifelse(daycountLH>=2 & daycountLH<=5, 1, 0), #added early lut
         LHgroup_midluteal_count = ifelse(midluteal_count==1, 1, 0),
         LHgroup_perimenstrual_count = ifelse(perimenstrual_count==1, 1, 0))


c2daily_LHsample <- c2daily_LHsample %>%
  rowwise() %>%
  mutate(sumdummy = sum(LHgroup_midfol_LH,
                        LHgroup_periov_LH,
                        LHgroup_perimenstrual_count,
                        LHgroup_midluteal_count, na.rm = T)) %>% 
  #if all phases=0, make the midluteal_LH variable NA instead of 0,
  #if either of the count variables are a 1, change midluteal_LH to 0 instead of NA to fill out structural set,
  #otherwise, keep as is
  mutate(LHgroup_midluteal_count = case_when(sumdummy==0 ~ as.numeric(NA), 
                                        sumdummy==1 & LHgroup_midfol_LH==1 ~ 0,
                                        sumdummy==1 & LHgroup_periov_LH==1 ~ 0,
                                        TRUE ~ LHgroup_midluteal_count), 
         #same for midfol
         LHgroup_midfol_LH = case_when(sumdummy==0 ~ as.numeric(NA), 
                                  sumdummy==1 & LHgroup_perimenstrual_count==1 ~ 0,
                                  sumdummy==1 & LHgroup_midluteal_count==1 ~ 0,
                                  TRUE ~ LHgroup_midfol_LH), 
          #same for periov
         LHgroup_periov_LH = case_when(sumdummy==0 ~ as.numeric(NA), 
                                  sumdummy==1 & LHgroup_perimenstrual_count==1 ~ 0,
                                  sumdummy==1 & LHgroup_midluteal_count==1 ~ 0,
                                  TRUE ~ LHgroup_periov_LH),
         #same for perimen
         LHgroup_perimenstrual_count = case_when(sumdummy==0 ~ as.numeric(NA), 
                                        sumdummy==1 & LHgroup_midfol_LH==1 ~ 0,
                                        sumdummy==1 & LHgroup_periov_LH==1 ~ 0,
                                        TRUE ~ LHgroup_perimenstrual_count))
```

#review
```{r}
#check overlap
c2daily_LHsample %>%
  group_by(LHgroup_periov_LH, LHgroup_midluteal_count, LHgroup_perimenstrual_count,LHgroup_midfol_LH) %>%
  summarize(n=n())

#review counting and phasing
c2daily_LHsample %>%
  select(id, daterated, posLHtoday, daycountLH,
         LHgroup_midfol_LH, LHgroup_periov_LH,
         firstdayofperiod, cycleday,
         LHgroup_midluteal_count, LHgroup_perimenstrual_count) %>%
  View()

```

#IF you want to save and review bin excel or separately
```{r}

c2daily_LHsample %>%
  select(id, daterated, posLHtoday, daycountLH,
         LHgroup_midfol_LH, LHgroup_periov_LH,
         firstdayofperiod, cycleday,
         LHgroup_midluteal_count, LHgroup_perimenstrual_count) %>%
  write_csv("2022-05-15 LH and MensesOnset Counting and Phasing.csv")

```


#review counting and phasing where there are OVERLAPS
```{r}

#periov and perimenstrual overlaps: n=3
c2daily_LHsample %>%
  filter(LHgroup_periov_LH==1) %>%
  filter(LHgroup_perimenstrual_count==1) %>%
  select(id, daterated, posLHtoday, daycountLH,
         menstrualbleeding,
         firstdayofperiod, cycleday,
         LHgroup_midfol_LH, LHgroup_periov_LH,
         LHgroup_midluteal_count, LHgroup_perimenstrual_count, cleartrialphase) %>%
  View()

#periov and midluteal overlaps: n=16
c2daily_LHsample %>%
  filter(LHgroup_midfol_LH==1) %>%
  filter(LHgroup_midluteal_count==1) %>%
  select(id, daterated, posLHtoday, daycountLH,
         menstrualbleeding,
         firstdayofperiod, cycleday,
         LHgroup_midfol_LH, LHgroup_periov_LH,
         LHgroup_midluteal_count, LHgroup_perimenstrual_count, cleartrialphase) %>%
  View()

#midluteal and midfol overlaps: n=4
c2daily_LHsample %>%
  filter(LHgroup_periov_LH==1) %>%
  filter(LHgroup_midluteal_count==1) %>%
  select(id, daterated, posLHtoday, daycountLH,
         menstrualbleeding,
         firstdayofperiod, cycleday,
         LHgroup_midfol_LH, LHgroup_periov_LH,
         LHgroup_midluteal_count, LHgroup_perimenstrual_count, cleartrialphase) %>%
  View()

#midfol and perimenstrual overlaps: n=10
c2daily_LHsample %>%
  filter(LHgroup_midfol_LH==1) %>%
  filter(LHgroup_perimenstrual_count==1) %>%
  select(id, daterated, posLHtoday, daycountLH,
         menstrualbleeding,
         firstdayofperiod, cycleday,
         LHgroup_midfol_LH, LHgroup_periov_LH,
         LHgroup_midluteal_count, LHgroup_perimenstrual_count, cleartrialphase) %>%
  View()

c2daily_LHsample %>%
  filter(LHgroup_periov_LH==1) %>%
  filter(LHgroup_perimenstrual_count==1) %>%
  select(id, daterated, posLHtoday, daycountLH,
         menstrualbleeding,
         firstdayofperiod, cycleday,
         LHgroup_midfol_LH, LHgroup_periov_LH,
         LHgroup_midluteal_count, LHgroup_perimenstrual_count) %>%
  View()
  

```

#create overrides
```{r}

c2daily_LHsample <- c2daily_LHsample %>%
  #if midluteal by backward count and periov by LH surge are both 1, keep periov and change midluteal to 0
  mutate(LHgroup_midluteal_count = case_when(LHgroup_periov_LH==1 ~ 0,
                                        TRUE ~ LHgroup_midluteal_count))

#check overlaps again
c2daily_LHsample %>%
  filter(id %in% c2_LHids) %>%
  group_by(LHgroup_periov_LH, LHgroup_midluteal_count, LHgroup_perimenstrual_count,LHgroup_midfol_LH) %>%
  summarize(n=n())

c2daily_LHsample <- c2daily_LHsample %>%
  #if perimenstrual by firstdayofperiod and midfol by pre-LH count are both 1, perimen wins, make midfol=0
  mutate(LHgroup_midfol_LH = case_when(LHgroup_perimenstrual_count==1 ~ 0,
                                        TRUE ~ LHgroup_midfol_LH))  


```


#Remove the cycles where data showed a 3-4 day luteal phase
-229: washout between C2+C3 (cleartrialphase 8-9)
-244: washout between C1+C2 (cleartrial phase 4-5)
```{r}

c2daily_LHsample <- c2daily_LHsample %>%
  filter(!(id==229 & cleartrialphase>7)) %>%
  filter(!(id==244 & (cleartrialphase==4 | cleartrialphase==5)))

```

# review again
```{r}
c1daily_LHsample %>%
  group_by(LHgroup_periov_LH, 
           LHgroup_midluteal_count, 
           LHgroup_perimenstrual_count, 
           LHgroup_midfol_LH) %>%
  summarize(n=n())
```

