---
title: "prep_3day-means"
author: "Jordan Barone"
date: "`r Sys.Date()`"
output: html_document
---

This code was used for calculating SI-related symptoms on the three days after a blood draw in the CLEAR1 dataset. We wanted to model the relationship between biomarkers and symptoms, and did a 3-day-post-visit average because:
1. There are many missing surveys on the actual lab visit days
2. We cannot use the in-lab survey, because it occurs BEFORE the blood draw

## library and data
```{r}

library(tidyverse)
library(haven)

df <- read.csv()

```


## create daily variables to go along with today's symptom, "next days symptom" and "symptom two days later"

```{r}

#create next day planning + next day symptoms
df <- df %>%
  arrange(id, date) %>%
  group_by(id) %>%
  mutate(nextday_ASIQplan = lead(ASIQ_planning),
         nextday_ASIQmean = lead(ASIQ_mean),
         twoday_ASIQplan = lead(ASIQ_planning, n=2),
         twoday_ASIQmean = lead(ASIQ_mean, n=2))


```


## create 72 hour means for SI variables
```{r SI variables, echo=TRUE}

#ASIQ_composite
df <- df %>%
  rowwise() %>%
  mutate(threeday_SI = mean(c(ASIQ_mean, nextday_ASIQmean, twoday_ASIQmean), na.rm=TRUE),
         threeday_dep = mean(c(DRSP1_depblue, nextday_DRSP1dep, twoday_DRSP1dep), na.rm=TRUE),
         threeday_anx = mean(c(DRSP4_anxious, nextday_DRSP4anx, twoday_DRSP4anx), na.rm=TRUE))

#log transform
narsad_full <- narsad_full %>%
  mutate(threeday_SI.log = log(threeday_SI),
         threeday_dep.log = log(threeday_dep),
         threeday_anx.log = log(threeday_anx))

#planning (0 or 1)
narsad_full <- narsad_full %>%
  rowwise() %>%
  mutate(threeday_plan = 
           case_when(ASIQ_planning==1 | nextday_ASIQplan==1 | twoday_ASIQplan==1 ~ 1,
                     TRUE ~ 0))


narsad_full <- narsad_full %>% 
  group_by(id) %>%
  mutate(SIMean_acrosslabs = mean(threeday_SI, na.rm=T),
         DEPMean_acrosslabs = mean(threeday_dep, na.rm=T),
         ANXMean_acrosslabs = mean(threeday_anx), na.rm=T)


```