---
title: "Calculating Symptom Cyclicity"
author: "Jordan Barone"
date: "Original date 8/9/2021; last edited/com"
output:
  pdf_document: default
  html_document: default
---

### Setup:
+ package import
```{r}
source("check_packages.R")
```

### file import
```{r}

data <- read_sav("~/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/....")

```

## create outcome list to run loop functions
+ just an example below, can obviously be any list of outcomes
```{r}

#create list of outcomes (only pre-registered)
outcomes <- c("SI_Mean",
              "DRSP1_depblue",
              "DRSP2_hopeless",
              "DRSP6_rejsens",
              "perceivedburden",
              "DRSP16_overwhelm",
              "DRSP4_anxious",
              "DRSP10_diffconc",
              "DRSP7_angirr",
              "DRSP11_lethtired", 
              "DRSP12_13_appoverate",
              "DRSP14_sleptmore",
              "DRSP15_troublesleep",
              "DRSP17_outofcontrol",
              "DRSP3_worthguilt",
              "DRSP9_lessintmot",
              "DRSP5_moodswings",
              "DRSP8_intconflict",
              "DRSP20_headache") %>% noquote()

outcomes.d <- outcomes %>% paste0(".d") %>% noquote()

```


## GRAND MEAN, PERSON MEAN, and PERSON DEVIATIONS: reference code
+ **person-mean calculation**
  + Dataset <- Dataset %>% group_by(id) %>%
   + mutate(impulsivity.m = mean(impulsivity, na.rm=T))

+ **person-centered deviation**
  + Dataset <- Dataset %>% group_by(id) %>%
    + rowwise() %>%
    + mutate(impulsivity.d = impulsivitytoday - impulsivity.m)
```{r}

#grand mean
dataset %>% summarize(minDRSP1 = round(min(DRSP1_depblue, na.rm = TRUE),2),
                      maxDRSP1 = round(max(DRSP1_depblue, na.rm = TRUE),2),
                      meanDRSP1 = round(mean(DRSP1_depblue, na.rm = TRUE),2),
                      sdDRSP1 = round(sd(DRSP1_depblue, na.rm = TRUE),2))

#person mean (just to see)
dataset %>% 
  group_by(id) %>% 
  summarise(DRSP1m = round(mean(DRSP1_depblue, na.rm = TRUE), digits = 2))


#save person mean as a trait level var
dataset <- dataset %>%
  group_by(id) %>%
  mutate(DRSP1m = mean(DRSP1_depblue, na.rm = TRUE)) 
# %>% select(id, DRSP1_depblue, DRSP1m)

#person mean centered, daily deviation
dataset <- dataset %>% group_by(id) %>% 
	rowwise() %>%
  mutate(DRSP1d = DRSP1_depblue - DRSP1.m) 

```

## functions to create person centered means and daily deviations
(full dataset of observations)
```{r}
#function to create person means
create.person.mean <- function(df, var, ...) {
  df %>%
    group_by(...) %>%
    mutate("{{var}}.m" := mean({{var}}, na.rm=T))
}
#function to create person deviations (note, must have person means already made)
create.deviation <- function(df, var, var.m) {
  df <- df %>%
    rowwise() %>%
    mutate("{{var}}.d" := {{var}} - {{var.m}})
}

```

## execute functions 
**Example on how to run these** \
+ to get a list of all person means: 
  + paste0(outcomelist,".m")
+ to run create.deviation() one at a time.
  + c2d <- create.deviation(c2d, wishsleep, wishsleep.m)
+ to run create.person.mean() one at a time.
  + c2d <- create.person.mean(c2d, ASIQ4_thoughtwhen, id)
```{r}
#execute for loop: run create.person.mean() on everything in "outcomelist"
for (i in outcomes) {
  dadtaset <- create.person.mean(dataset, !!sym({{i}}), id)
}
#execute for loop: run create.deviation() on everything in list
for (i in outcomes) {
  dataset <- create.deviation(dataset, !!sym({{i}}), !!sym(paste0({{i}}, ".m")))
}
```

## CENTER - all affect vars
```{r}

#grand mean
overall_summary <- etoh_daily %>% 
  summarize(minDRSP1 = round(min(DRSP1_depblue, na.rm = TRUE),2),
                      maxDRSP1 = round(max(DRSP1_depblue, na.rm = TRUE),2),
                      meanDRSP1 = round(mean(DRSP1_depblue, na.rm = TRUE),2),
                      sdDRSP1 = round(sd(DRSP1_depblue, na.rm = TRUE),2),
                      minDRSP2 = round(min(DRSP2_hopeless, na.rm = TRUE),2),
                      maxDRSP2 = round(max(DRSP2_hopeless, na.rm = TRUE),2),
                      meanDRSP2 = round(mean(DRSP2_hopeless, na.rm = TRUE),2),
                      sdDRSP2 = round(sd(DRSP2_hopeless, na.rm = TRUE),2),
                      minDRSP3 = round(min(DRSP3_worthguilt, na.rm = TRUE),2),
                      maxDRSP3 = round(max(DRSP3_worthguilt, na.rm = TRUE),2),
                      meanDRSP3 = round(mean(DRSP3_worthguilt, na.rm = TRUE),2),
                      sdDRSP3 = round(sd(DRSP4_anxious, na.rm = TRUE),2),
                      minDRSP4 = round(min(DRSP4_anxious, na.rm = TRUE),2),
                      maxDRSP4 = round(max(DRSP4_anxious, na.rm = TRUE),2),
                      meanDRSP4 = round(mean(DRSP5_moodswings, na.rm = TRUE),2),
                      sdDRSP4 = round(sd(DRSP5_moodswings, na.rm = TRUE),2),
                      minDRSP5 = round(min(DRSP5_moodswings, na.rm = TRUE),2),
                      maxDRSP5 = round(max(DRSP5_moodswings, na.rm = TRUE),2),
                      meanDRSP5 = round(mean(DRSP5_moodswings, na.rm = TRUE),2),
                      sdDRSP5 = round(sd(DRSP5_moodswings, na.rm = TRUE),2),
                      minDRSP6 = round(min(DRSP6_rejsens, na.rm = TRUE),2),
                      maxDRSP6 = round(max(DRSP6_rejsens, na.rm = TRUE),2),
                      meanDRSP6 = round(mean(DRSP6_rejsens, na.rm = TRUE),2),
                      sdDRSP6 = round(sd(DRSP6_rejsens, na.rm = TRUE),2),
                      minDRSP7 = round(min(DRSP7_angirr, na.rm = TRUE),2),
                      maxDRSP7 = round(max(DRSP7_angirr, na.rm = TRUE),2),
                      meanDRSP7 = round(mean(DRSP7_angirr, na.rm = TRUE),2),
                      sdDRSP7 = round(sd(DRSP7_angirr, na.rm = TRUE),2),
                      minDRSP8 = round(min(DRSP8_intconflict, na.rm = TRUE),2),
                      maxDRSP8 = round(max(DRSP8_intconflict, na.rm = TRUE),2),
                      meanDRSP8 = round(mean(DRSP8_intconflict, na.rm = TRUE),2),
                      sdDRSP8 = round(sd(DRSP8_intconflict, na.rm = TRUE),2)) %>%
  t()

```


## CREATE PERSON-LEVEL AVG CYCLICITY: CPASS THRESHOLDS
-( average POSTmenstrual ratings MINUS average PREmenstrual ratings ) / range of scale * 100
-range of scale is calculated based on the highest and lowest score they gave ANY DRSP item (example: if hopelessness is only ever rated 1-3, but anxiety goes 1-6, their range of scale is always 5)
*premenstrual*  days (-7) to (-1)
*postmenstrual* days (4) to (10)
```{r}
#create binary vars for premenstrual
etoh_daily <- etoh_daily %>% mutate(CPASS_pre = ifelse(cycleday >= -7 & cycleday <= -1, 1, 0),
                   CPASS_post = ifelse(cycleday >= 4 & cycleday <= 10, 1, 0))

#check that nobody has <=3 pre or post menstrual days
etoh_daily %>% group_by(id, CPASS_pre, CPASS_post) %>% 
  filter(CPASS_pre==1 | CPASS_post==1) %>%
  summarize(n=n()) %>% 
  arrange(n)

```


```{r}

cpasstemp1 <- etoh_daily %>% group_by(id) %>%
  summarize(DRSPmin= min(DRSP1_depblue,
                     DRSP2_hopeless,
                     DRSP3_worthguilt,
                     DRSP4_anxious,
                     DRSP5_moodswings,
                     DRSP6_rejsens,
                     DRSP7_angirr,
                     DRSP8_intconflict, na.rm = TRUE),
            DRSPmax= max(DRSP1_depblue,
                     DRSP2_hopeless,
                     DRSP3_worthguilt,
                     DRSP4_anxious,
                     DRSP5_moodswings,
                     DRSP6_rejsens,
                     DRSP7_angirr,
                     DRSP8_intconflict, na.rm = TRUE))

cpasstemp1 <- cpasstemp1 %>% mutate(ROS = DRSPmax -1)


#create a PREmenstrual mean and a POSTmenstrual mean for each person
CPASS_pre <- etoh_daily %>% filter(CPASS_pre==1) %>% group_by(id) %>%
  summarize(cpasspre_meanDRSP1 = mean(DRSP1_depblue, na.rm= TRUE),
            cpasspre_meanDRSP2 = mean(DRSP2_hopeless, na.rm= TRUE),
            cpasspre_meanDRSP3 = mean(DRSP3_worthguilt, na.rm= TRUE),
            cpasspre_meanDRSP4 = mean(DRSP4_anxious, na.rm= TRUE),
            cpasspre_meanDRSP5 = mean(DRSP5_moodswings, na.rm= TRUE),
            cpasspre_meanDRSP6 = mean(DRSP6_rejsens, na.rm= TRUE),
            cpasspre_meanDRSP7 = mean(DRSP7_angirr, na.rm= TRUE),
            cpasspre_meanDRSP8 = mean(DRSP8_intconflict, na.rm= TRUE))

CPASS_post <- etoh_daily %>% filter(CPASS_post==1) %>% group_by(id) %>%
  summarize(cpasspost_meanDRSP1 = mean(DRSP1_depblue, na.rm= TRUE),
            cpasspost_meanDRSP2 = mean(DRSP2_hopeless, na.rm= TRUE),
            cpasspost_meanDRSP3 = mean(DRSP3_worthguilt, na.rm= TRUE),
            cpasspost_meanDRSP4 = mean(DRSP4_anxious, na.rm= TRUE),
            cpasspost_meanDRSP5 = mean(DRSP5_moodswings, na.rm= TRUE),
            cpasspost_meanDRSP6 = mean(DRSP6_rejsens, na.rm= TRUE),
            cpasspost_meanDRSP7 = mean(DRSP7_angirr, na.rm= TRUE),
            cpasspost_meanDRSP8 = mean(DRSP8_intconflict, na.rm= TRUE))

cpass_cyclicity <- left_join(cpasstemp1, CPASS_pre, by="id") %>%
  left_join(., CPASS_post, by="id") %>%
  rowwise() %>%
  mutate(DRSP1_cpass = ((cpasspre_meanDRSP1 - cpasspost_meanDRSP1)/ROS)*100,
         DRSP2_cpass = ((cpasspre_meanDRSP2 - cpasspost_meanDRSP2)/ROS)*100,
         DRSP3_cpass = ((cpasspre_meanDRSP3 - cpasspost_meanDRSP3)/ROS)*100,
         DRSP4_cpass = ((cpasspre_meanDRSP4 - cpasspost_meanDRSP4)/ROS)*100,
         DRSP5_cpass = ((cpasspre_meanDRSP5 - cpasspost_meanDRSP5)/ROS)*100,
         DRSP6_cpass = ((cpasspre_meanDRSP6 - cpasspost_meanDRSP6)/ROS)*100,
         DRSP7_cpass = ((cpasspre_meanDRSP7 - cpasspost_meanDRSP7)/ROS)*100,
         DRSP8_cpass = ((cpasspre_meanDRSP8 - cpasspost_meanDRSP8)/ROS)*100)
  
plot(cpass_cyclicity$DRSP1_cpass) #depblue
plot(cpass_cyclicity$DRSP2_cpass) #hopeless
plot(cpass_cyclicity$DRSP3_cpass) #worthguilt
plot(cpass_cyclicity$DRSP4_cpass) #anxious
plot(cpass_cyclicity$DRSP5_cpass) #moodswings
plot(cpass_cyclicity$DRSP6_cpass) #rejsens
plot(cpass_cyclicity$DRSP7_cpass) #angirr
plot(cpass_cyclicity$DRSP8_cpass) #intconflict

hist(cpass_cyclicity$DRSP1_cpass)
hist(cpass_cyclicity$DRSP2_cpass)
hist(cpass_cyclicity$DRSP3_cpass)
hist(cpass_cyclicity$DRSP4_cpass)
hist(cpass_cyclicity$DRSP5_cpass)
hist(cpass_cyclicity$DRSP6_cpass)
hist(cpass_cyclicity$DRSP7_cpass)
hist(cpass_cyclicity$DRSP8_cpass)

##CREATE CATEGORICAL THRESHOLDS
cpass_cyclicity <- cpass_cyclicity %>%
  mutate(over30 = ifelse(any(DRSP1_cpass:DRSP8_cpass >= 30), 1, 0),
         underneg30 = ifelse(any(DRSP1_cpass:DRSP8_cpass <= -30), 1, 0))

#count of participants who had PREmenstrual over POSTmenstrual ratings > 30% change
cpass_cyclicity %>%
  group_by(over30) %>%
  summarize(pre_high = n())

#count of participants who had POSTmenstrual ratings over PREmenstrual ratings > 30% change
cpass_cyclicity %>%
  group_by(underneg30) %>%
  summarize(pre_low = n())


```


archive??? below sections calculate cyclicity based on HTSMC paper phases, not cpass phases

## CREATE PERSON-LEVEL AVG CYCLICITY
FIRST confirm there is at least 3 days per phase, per person
-221: only 2 days in perimenstrual phase
-212, 217: only 3 days in either perimenstrual or midfol
```{r}
#FIRST confirm there is at least 3 days per phase, per person
etoh_daily %>% group_by(id, perimenstrual_count, midfol_count) %>% 
  filter(perimenstrual_count==1 | midfol_count==1) %>%
  summarize(n=n()) %>% 
  arrange(n)
#remove: 221, 212, 217

etoh_daily_premen_v_midfol <- etoh_daily %>% 
  filter(id!=221) %>%
  filter(id!=212) %>%
  filter(id!=217)


```

## AVG CYCLICITY
-( average perimenstrual ratings MINUS average midfollicular ratings ) / range of scale
-range of scale is calculated based on the highest and lowest score they gave ANY DRSP item (example: if hopelessness is only ever rated 1-3, but anxiety goes 1-6, their range of scale is always 5)
-combine perimenstrual and midfollicular from as many cycles as we have from that person
-use RAW ratings (not centered)
```{r}

#create grouped dataframe of each ID's min and max DRSP rating
temp1 <- etoh_daily_premen_v_midfol %>% group_by(id) %>% #first, group by id
  summarize(DRSPmin= min(DRSP1_depblue, 
 # min() is the function, then thelist tells R "get the MINIMUM of everything I've listed out#
                     DRSP2_hopeless,
                     DRSP3_worthguilt,
                     DRSP4_anxious,
                     DRSP5_moodswings,
                     DRSP6_rejsens,
                     DRSP7_angirr,
                     DRSP8_intconflict, na.rm = TRUE),
            DRSPmax= max(DRSP1_depblue, #same for max
                     DRSP2_hopeless,
                     DRSP3_worthguilt,
                     DRSP4_anxious,
                     DRSP5_moodswings,
                     DRSP6_rejsens,
                     DRSP7_angirr,
                     DRSP8_intconflict, na.rm = TRUE))

#the ROS (range of scale) is the MAX minus 1
temp1 <- temp1 %>% mutate(ROS = DRSPmax -1)

#create a perimenstrual mean and a midfollicular mean for each person
temp2 <- etoh_daily_premen_v_midfol %>% filter(perimenstrual_count==1) %>% group_by(id) %>%
  summarize(perimen_meanDRSP1 = mean(DRSP1_depblue, na.rm= TRUE),
            perimen_meanDRSP2 = mean(DRSP2_hopeless, na.rm= TRUE),
            perimen_meanDRSP3 = mean(DRSP3_worthguilt, na.rm= TRUE),
            perimen_meanDRSP4 = mean(DRSP4_anxious, na.rm= TRUE),
            perimen_meanDRSP5 = mean(DRSP5_moodswings, na.rm= TRUE),
            perimen_meanDRSP6 = mean(DRSP6_rejsens, na.rm= TRUE),
            perimen_meanDRSP7 = mean(DRSP7_angirr, na.rm= TRUE),
            perimen_meanDRSP8 = mean(DRSP8_intconflict, na.rm= TRUE))

temp3 <- etoh_daily_premen_v_midfol %>% filter(midfol_count==1) %>% group_by(id) %>%
  summarize(midfol_meanDRSP1 = mean(DRSP1_depblue, na.rm= TRUE),
            midfol_meanDRSP2 = mean(DRSP2_hopeless, na.rm= TRUE),
            midfol_meanDRSP3 = mean(DRSP3_worthguilt, na.rm= TRUE),
            midfol_meanDRSP4 = mean(DRSP4_anxious, na.rm= TRUE),
            midfol_meanDRSP5 = mean(DRSP5_moodswings, na.rm= TRUE),
            midfol_meanDRSP6 = mean(DRSP6_rejsens, na.rm= TRUE),
            midfol_meanDRSP7 = mean(DRSP7_angirr, na.rm= TRUE),
            midfol_meanDRSP8 = mean(DRSP8_intconflict, na.rm= TRUE))

cyclicity <- left_join(temp1, temp2, by="id") %>%
  left_join(., temp3, by="id") %>%
  rowwise() %>%
  mutate(DRSP1_cyc = ((perimen_meanDRSP1 - midfol_meanDRSP1)/ROS)*100,
         DRSP2_cyc = ((perimen_meanDRSP2 - midfol_meanDRSP2)/ROS)*100,
         DRSP3_cyc = ((perimen_meanDRSP3 - midfol_meanDRSP3)/ROS)*100,
         DRSP4_cyc = ((perimen_meanDRSP4 - midfol_meanDRSP4)/ROS)*100,
         DRSP5_cyc = ((perimen_meanDRSP5 - midfol_meanDRSP5)/ROS)*100,
         DRSP6_cyc = ((perimen_meanDRSP6 - midfol_meanDRSP6)/ROS)*100,
         DRSP7_cyc = ((perimen_meanDRSP7 - midfol_meanDRSP7)/ROS)*100,
         DRSP8_cyc = ((perimen_meanDRSP8 - midfol_meanDRSP8)/ROS)*100)
  
plot(cyclicity$DRSP1_cyc) #depblue
plot(cyclicity$DRSP2_cyc) #hopeless
plot(cyclicity$DRSP3_cyc) #worthguilt
plot(cyclicity$DRSP4_cyc) #anxious
plot(cyclicity$DRSP5_cyc) #moodswings
plot(cyclicity$DRSP6_cyc) #rejsens
plot(cyclicity$DRSP7_cyc) #angirr
plot(cyclicity$DRSP8_cyc) #intconflict

hist(cyclicity$DRSP1_cyc)
hist(cyclicity$DRSP2_cyc)
hist(cyclicity$DRSP3_cyc)
hist(cyclicity$DRSP4_cyc)
hist(cyclicity$DRSP5_cyc)
hist(cyclicity$DRSP6_cyc)
hist(cyclicity$DRSP7_cyc)
hist(cyclicity$DRSP8_cyc)


##CREATE CATEGORICAL THRESHOLDS
cyclicity <- cyclicity %>%
  mutate(over30 = ifelse(any(DRSP1_cyc:DRSP8_cyc >= 30), 1, 0),
         underneg30 = ifelse(any(DRSP1_cyc:DRSP8_cyc <= -30), 1, 0))

#count of participants who had perimenstrual over midfollicular ratings > 30% change
cyclicity %>%
  group_by(over30) %>%
  summarize(perimen_high = n())

#count of participants who had midfollicular ratings over perimenstrual ratings > 30% change
cyclicity %>%
  group_by(underneg30) %>%
  summarize(perimen_low = n())


```