---
title: "CYCLE variables"
author: "Jordan Barone"
date: "2022-09-02"
output: html_document
---

*NOTE: it is absolutely essential that date duplicates are cleaned before adding cycle variables.*

# add cycle day
```{r cycleday}
#if menstrualbleeding = 0.5 (spotting) or 1 (bleeding) and firstdayofperiod = 1, then cycleday = 1
Clear2DailyData$cycleday <- ifelse(Clear2DailyData$menstrualbleeding >= 0.5 & Clear2DailyData$firstdayofperiod == 1, 1, 0)
#convert all NAs in cycleday to 0
Clear2DailyData$cycleday[is.na(Clear2DailyData$cycleday)] <- 0

#Fill in missing days by ID 
Clear2DailyData <- Clear2DailyData[!is.na(Clear2DailyData$daterated), ]
Clear2DailyData <- Clear2DailyData %>% group_by(id) %>%
  complete(daterated = seq.Date(min(daterated), max(daterated), by = "day"))
Clear2DailyData <- cbind(Clear2DailyData, 
                         year = year(Clear2DailyData$daterated), 
                         month = month(Clear2DailyData$daterated), 
                         day = day(Clear2DailyData$daterated)) 
Clear2DailyData <- Clear2DailyData[with(Clear2DailyData,order(as.numeric(id), 
                                                              year, month, day, daterated), cycleday),]
Clear2DailyData$cycleday[is.na(Clear2DailyData$cycleday)] <- 0

Clear2DailyData <- Clear2DailyData[with(Clear2DailyData,order(id, 
                                                              year, 
                                                              month, 
                                                              day, 
                                                              daterated),cycleday),]

#make A the variable where menstrualbleeding starts AND period starts
Clear2DailyData$A <- Clear2DailyData$cycleday

cycleCount <- function(x) {
  #Get the index of 1
  inds <- which(x == 1)
  if(!length(inds)) return(0) 
  #Create a sequence with that index as 0
  num <- lapply(inds, function(i) {
    num <- seq_along(x) - i
    #Add 1 to values greater than equal to 0
    num[num >= 0] <- num[num >= 0] + 1
    num[num < -15 | num > 10] <- NA #jb changed days here 8/24/21
    num
  })
  #Select the first non-NA values from the sequence
  do.call(coalesce, num)
}

Clear2DailyData <- Clear2DailyData %>% group_by(id) %>% 
  mutate(cycleday = cycleCount(A))

Clear2DailyData <- Clear2DailyData %>% select(!A)
```

#ways to view and check that the cycle day code looks right
```{r}

#View in a separate window and manually scroll through
Clear2DailyData %>% select(id, daterated, firstdayofperiod, cycleday, posLHtoday, cycleday) %>% View()

#how many observations do you have for each cycle day?
Clear2DailyData %>% group_by(cycleday) %>% summarize(n=n())

#do you have any observations that got assigned a cycle day of 0? this is wrong, and you should check who that happened for
Clear2DailyData %>% filter(cycleday==0) %>% pull(id)


```

### add cycle phase
 - goal is to have 4 dummy code variables with 0, 1, and NA for each of 4 phases
 - for ref, RAW clear1 has this code: 1=fol, 2=ov, 3=midlut, 4=PREmenstrual [this will be removed]
 - for future *visualizing*, it is also helpful to have one variable with all of the phases - this will NOT BE USED for analysis
```{r}
##ADD cycle phase based on "how to study" paper

Clear2DailyData <- Clear2DailyData %>% 
  mutate(midluteal_count = ifelse(cycleday >= -9 & cycleday <= -5, 1, 0),
         perimenstrual_count = ifelse(cycleday >=-3 & cycleday <=2, 1, 0),
         midfol_count = ifelse(cycleday >=4 & cycleday <=7, 1, 0),
         periov_count = ifelse(cycleday >= -15 & cycleday <= -12, 1, 0))

#add a non-dummy coded variable with all of cycle phase info above, which will be useful for future visualizations and summaries
Clear2DailyData <- Clear2DailyData %>% mutate(cyclephase = case_when(periov_count==1 ~ 1,
                                                     midluteal_count==1 ~ 2,
                                                     perimenstrual_count==1 ~ 3,
                                                     midfol_count==1 ~ 4,
                                                     TRUE ~ as.numeric(NA)))

```

