---
title: "prep_how_many_cycles_per_id?"
author: "Anisha Nagpal"
date: "`r Sys.Date()`"
output: html_document
---

```{r}
library(dplyr)
library(haven) #Use to import .sav files 
library(tidyr)
library(stringr)
library(devtools)
library(tidyverse)
```

Use the code below to create a new variable called 'cycle' which tells you which cycle a particular row of data is contributing to, within a person. In order to run the code below you must have the variables 'id' and 'cycleday' in your dataset. This prep code is necessary to run the R cpass package. The code below assumes your dataframe is called 'dat'. You can change the name of 'dat' to match whatever your dataframe is. Please reach out to Anisha if you have any questions! 



```{r}
dat <- dat %>% group_by(id) %>% arrange(date_rated, .by_group = TRUE)

# create new cycle variable
dat$cycle <- NA

# iterate over id groups
for (id in unique(dat$id)) {
  cycle_count <- 1
  stretch_count <- 1
  stretch_flag <- FALSE
  prev_cycleday <- NA
  last_cycle <- NA
  for (i in which(dat$id == id)) {
    # skip rows with NA cycleday
    if (is.na(dat$cycleday[i])) {
      dat$cycle[i] <- NA
      stretch_flag <- FALSE
      prev_cycleday <- NA
      next
    }
    # increment cycle_count if current row starts a new stretch
    if (!stretch_flag && dat$cycleday[i] %in% c(-15, -14, -13, -12, -11, -10, 
                                                  -9, -8, -7, -6, -5, -4, 
                                                  -3, -2, -1, 1, 2, 3, 4, 
                                                  5, 6, 7, 8, 9, 10, -1, 1) && 
        (is.na(prev_cycleday) || !(prev_cycleday %in% c(-15, -14, -13, -12, -11, -10, 
                                                  -9, -8, -7, -6, -5, -4, 
                                                  -3, -2, -1, 1, 2, 3, 4, 
                                                  5, 6, 7, 8, 9, 10, -1, 1)))) {
      # increment cycle by 1 if there is a last cycle value within the same id group
      if (!is.na(last_cycle)) {
        cycle_count <- last_cycle + 1
      }
      dat$cycle[i] <- cycle_count
      stretch_flag <- TRUE
    }
    # update cycle variable within a stretch
    else if (stretch_flag) {
      dat$cycle[i] <- cycle_count
      # increment stretch_count if current row is the end of a stretch
      if (!dat$cycleday[i] %in% c(-15, -14, -13, -12, -11, -10, -9, -8, 
                                   -7, -6, -5, -4, -3, -2, -1, 1, 2, 
                                   3, 4, 5, 6, 7, 8, 9, 10, -1, 1)) {
        stretch_count <- stretch_count + 1
        cycle_count <- cycle_count + 1
        stretch_flag <- FALSE
      }
    }
    prev_cycleday <- dat$cycleday[i]
  last_cycle <- dat$cycle[i]
  }
}


dat %>%
  select(id, cycleday, StartPeriod, cycle) #this is to just check how id, cycleday, StarttPeriod, and cycle all line up. StartPeriod is the same as firstdayperiod. 
```

